Efficient GPU implementation of multicomponent and multiphase Lattice Boltzmann models.

### 2010-04-20 paper.tex
Added the one component description with Julia's model. Still cannot make the model working with the Shan-Chen potential even with smoothing.

\documentclass[mathpazo,sort,numbers]{cicp}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{color}
%\usepackage[sort]{cite}
%\usepackage[sort&compress,authoryear,round]{natbib}

@@

This was quickly noticed by the CFD community and in the last few years the method was
implemented on NVIDIA GPUs and applied to many different problems, such as calculation of
the drag coefficient for the two-dimensional flow through cylinders \cite{tolke-twod} and
for a moving sphere in a three-dimensional channel \cite{tolke-GPU}.  In the majority of these
problems only fully local, single-fluid LB models were used.  Only very recently some
work on more complex models with NN interactions has been done {\color{red} cite Succi's paper},
but using a code ported from a CPU version, and not optimized specifically for the GPU
environment.
%%
% To the authors' best knowledge only simple
% hydrodynamics problems GPUs implementation is thoroughly described in the literature.
%%
% This is actually not true anymore -- see the Succi's PRE paper about LBM on GPUs
The aim of this paper is to present efficient GPU-optimized algorithms for LB models
with NN interaction terms.  Such models are abundant and of enormous practical importance.
They include the Shan-Chen model, free-energy binary fluid models and higher-order
boundary conditions for fluid flows.  In the following sections, we first present
the details of the Shan-Chen and free-energy LB models.  The reader
is then briefly introduced to the main features of the CUDA environment
that we use to program the GPUs, which is followed by a thorough description
of the GPU algorithms and an analysis of their efficiency.  In section \ref{sec:bechmark}, the two
binary fluid models are compared in terms of performance and accuracy.
We conclude the paper with a summary of the main findings.

@@

@@

{\color{red} Michal, need your help here. Which parameters have you used for simulation. Please check above.}
We take equilibrium distribution function for the binary-liquid gas model. In general, it can be presented in the following form:

\subsection{Lattice Boltzmann implementation}
For all distribution functions the collision and streaming step are performed:
\begin{equation*}
\begin{aligned}
f_i^{\star}=(1-\omega)f_i+f_i^{eq}\\
f_i(\bm{x}+\bm{c_i})=f_i^{\star}. 
\end{aligned}
\end{equation*}

@@

The slowest kind of memory is the host memory accessible only to the CPU.  This memory,
while potentially very large in size, cannot be accessed directly by the GPU, so in order
for data stored there to be useful during a GPU simulation, it has to be copied to the
global device memory.  Such a data transfer goes over the PCI-e bus, which has a maximum
throughput of about 6~GB/s.

Next in line is the global device memory of the GPU, which can currently span several
gigabytes and which has a bandwidth of about 100~GB/s.  Global memory accesses are however
high-latency operations, taking several hundred clock cycles of the GPU to complete {\color{red} Cite NVIDIA documentation here}

>>>>

\documentclass[mathpazo]{cicp}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{color}
%\usepackage[sort]{cite}
%\usepackage[sort&compress,authoryear,round]{natbib}

@@

This was quickly noticed by the CFD community and in the last few years the method was
implemented on NVIDIA GPUs and applied to many different problems, such as calculation of
the drag coefficient for the two-dimensional flow through cylinders \cite{tolke-twod} and
for a moving sphere in a three-dimensional channel \cite{tolke-GPU}.  In the majority of these
problems only fully local, single-fluid LB models were used.  Only very recently some
work on more complex models with NN interactions has been done {\color{red} cite Succi's paper},
but using a code ported from a CPU version, and not optimized specifically for the GPU
environment.
%%
% To the authors' best knowledge only simple
% hydrodynamics problems GPUs implementation is thoroughly described in the literature.
%%
% This is actually not true anymore -- see the Succi's PRE paper about LBM on GPUs
The aim of this paper is to present efficient GPU-optimized algorithms for LB models
with NN interaction terms.  Such models are abundant and of enormous practical importance.
They include the Shan-Chen model, free-energy binary fluid models and higher-order
boundary conditions for fluid flows.  In the following sections, we first present
the details of the Shan-Chen and free-energy LB models.  The reader
is then briefly introduced to the main features of the CUDA environment
that we use to program the GPUs, which is followed by a thorough description
of the GPU algorithms and an analysis of their efficiency.  In section \ref{sec:benchmark}, the two
binary fluid models are compared in terms of performance and accuracy.
We conclude the paper with a summary of the main findings.

@@

The lattice Boltzmann implementation in a standard view:
\begin{equation}
\label{standard:implementation}
\begin{aligned}
&f_i^{*}(\bm{x},t)=\omega f_i^{eq}(\bm{x},t)-(1-\omega) f_i(\bm{x},t)\\
&f_i(\bm{x}+\bm{c_i},t+1)=f_i^{*}(\bm{x},t). 
\end{aligned}
\end{equation}

@@

{\color{red} Michal, need your help here. Which parameters have you used for simulation. Please check above.}
We take equilibrium distribution function for the binary-liquid gas model. In general, it can be presented in the following form:
\subsection{Free-energy model}
The free-energy or pressure model \cite{swift} allows the incorporation of the any equation of state to the lattice Boltzmann equation. The lattice Boltzmann scheme is implemented in a standard form (\ref{standard:implementation}), but the equilibrium function is defined as:
\begin{equation}
\begin{aligned}
f_i^{eq}=&w_i \biggl(p_0(\rho)-3 k\rho\Delta\rho+\rho \frac{c_{i\alpha}u_{\alpha}}{3}+\rho \frac{9 Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2}\biggr)\\
&+k (w_{ixx} \partial_x \rho \partial_x \rho + w_{ixy} \partial_x \rho \partial_y \rho + w_{iyy} \partial_y \rho \partial_y \rho),
\end{aligned}
\end{equation}
where $w_{ixx}=\{\dfrac{1}{3},\dfrac{1}{3},-\dfrac{1}{6},-\dfrac{1}{6},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24}\}$, $w_{iyy}=\{\dfrac{1}{3},\dfrac{1}{3},-\dfrac{1}{6},-\dfrac{1}{6},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24}\}$ and $w_{ixy}=\{0,0,0,0,\dfrac{1}{4},\dfrac{1}{4},-\dfrac{1}{4},-\dfrac{1}{4}\}$. The laplacian and the gradient are taken by the numerical stencils:
\begin{equation}
\Delta=\frac{1}{6}
\begin{pmatrix}
1&4&1\\
4&-20&4\\
1&4&1\\
\end{pmatrix}\text{ and }
\partial_x=\frac{1}{12}
\begin{pmatrix}
-1&0&1\\
-4&0&4\\
-1&0&1\\
\end{pmatrix}.
\end{equation}

We implement in terms of convergence,accuracy and performance this model with the Shan-Chen potential, mainly $p_0=\dfrac{\rho}{3}+G\dfrac{\psi^2(\rho)}{6}$.

\subsection{Lattice Boltzmann implementation}
For all distribution functions the collision and streaming step are performed:
\begin{equation*}
\begin{aligned}
&f_i^{\star}=(1-\omega)f_i+f_i^{eq}\\
&f_i(\bm{x}+\bm{c_i})=f_i^{\star}. 
\end{aligned}
\end{equation*}

@@

The slowest kind of memory is the host memory accessible only to the CPU.  This memory,
while potentially very large in size, cannot be accessed directly by the GPU, so in order
for data stored there to be useful during a GPU simulation, it has to be copied to the
global device memory.  Such a data transfer goes over the PCI-e bus, which has a maximum
throughput of about 6~GB/s.
% max 3.2 GB/s one way, measured on a 4xGTX285 system

Next in line is the global device memory of the GPU, which can currently span several
gigabytes and which has a bandwidth of about 100~GB/s.  Global memory accesses are however
high-latency operations, taking several hundred clock cycles of the GPU to complete {\color{red} Cite NVIDIA documentation here}
% 131.9 GB/s for GTX 280, see NVIDIA CUDA Best Practices Guide, v3.0
